---
title: "R Notebook"
output: 
  md_document:
    variant: markdown_github
---

# netprobitFE

R package that implements the methods developed in the paper 

Dzemski, A: "An empirical model of dyadic link formation in a network with unobserved heterogeneity", Review of Economics and Statistics, forthcoming.

## Installation

The easiest way to install the package is to use the `devtools` package.
Make sure that you have the `devtools` package installed and type:

```R
devtools::install_github("adzemski/netprobitFE")
```

## Example

If you are interested in using this package you probably have your own network data that you want to analyze.
For this demonstration, we start without any data and we will have to make some data up. 
Let's use the package to simulate an example network with $N=50$ agents.

```{r paged.print=FALSE}
library(netprobitFE)
# first define parameters of simulation
sim_design <- define_sim_design_jochmans(N = 50, theta = 1, rho = 0.5, type_C_N = "loglog")
# generate a network with these parameters
set.seed(123)
dyadic_data <- as.data.frame(draw_network_jochmans_2018(sim_design)$links)[, c("i", "j", "X1", "Y")]
head(dyadic_data, n = 10)
```

Our analysis of the simulated network data in `dyadic_data` starts here.
We start by telling the package about what data we are going to use and what model we are going to estimate.

```{r}
probit_model <- define_model(dyadic_data, X_names = "X1", col_sender = "i", col_receiver = "j",
                                          col_outcome = "Y")
```

The data is expected to be dataframe that contains the link-specific covariates, the link outcome ($0$ for no link, $1$ for a link), and the sender identifier and the receiver identifier. The unit identifier that specifies the sender and the receiver of a link can be a number or a string. 
The data does not have to be ordered in a particular way but it has to be *complete* in the sense that it has to contain one row for every possible link (with $N$ agents there are $N(N-1)$ possible links). To specify more than one link-specific covariate use a vector. For example, if there are three covariates `X1`, `X2` and `X3` then replace `X_names = "X1"` by `X_names = c("X1, "X2", "X3")`.

We now estimate the homophily parameter $\theta$ and the reciprocity parameter $\rho$ using the two-stage procedure 
described in Section 2.2 of the paper.

```{r}
# stage 1
fit_ML <- MLE_stage1(probit_model)
fit_ML$theta_hat_MLE
# stage 2
fit_ML <- MLE_stage2(fit_ML)
fit_ML$rho_hat_MLE
```

The procedure `ttest_theta` computes a bias-corrected estimate of $\theta$ and conducts $t$-tests that are robust to the incidental parameter problem. For example we can test significance, i.e., test $\theta$ against zero:

```{r}
fit_ttest_theta <- ttest_theta(fit_ML, theta0 = 0)
fit_ttest_theta[c("theta_hat_bias_corr", "tstat_theta", "pval_theta")]
```

The test rejects at the 5% level. This means that `X1` is significant at the 5% level. Note that the bias-corrected estimate is indeed closer to the true value ($\theta = 1$) than the uncorrected ML estimate. We can also test against the true value of $\theta$:

```{r}
ttest_theta(fit_ML, theta0 = 1)[c("theta_hat_bias_corr", "tstat_theta", "pval_theta")]
```

Unsurprisingly, this test does not reject. 

The procedure `ttest_rho` computes a bias-corrected estimate of $\rho$ and conducts $t$-tests that are robust to the incidental parameter problem. For example, we can test significance: 

```{r}
ttest_rho(fit_ML, fit_ttest_theta, rho0 = 0)[c("rho_hat_bias_corr", "pval_rho")]
```

The test rejects at the 5% level. This means that the within dyad shocks are correlated (95% confidence). Note that the bias-corrected estimate is indeed closer to the true value ($\rho = 0.5$) than the uncorrected ML estimate. We can also test against the true value of $\theta$:
 
```{r}
ttest_rho(fit_ML, fit_ttest_theta, rho0 = 0.5)[c("rho_hat_bias_corr", "pval_rho")]
```

This test rejects at the 5% level. 

Finally, we test model specification using the transitivity test described in Section 5 of the paper. We set the nominal size of the test to $\alpha = 0.5$ and turn bootstrap inference on (as recommended in the paper). Here we set the number of bootstrap iterations (`bs_iterations`) to $200$. The bootstrap is fairly expensive and this will take a while. On my computer the specification test takes about 37 seconds to execute. In practice, I recommend to use a much larger number of bootstrap iterations. 

```{r}
triangle_test(fit_ML, fit_ttest_theta, bs_inference = "double", 
              bs_iterations = 200, alpha = 0.05)[c("excess_trans", "excess_trans_bias_corr", "teststat",
                                                   "pval_analytic", "rej_bs_perc_t")]
```

The statistic `excess_trans_bias` is the naive excess transitivity measure $T_N$ from Section 5.2, the statistic `excess_trans_bias_corr` is the same statistic with analytic bias correction based on Theorem 3. Finally, `teststat` is the studentized robust test statistic given in equation (5.3). The percentile bootstrap does not reject (`rej_bs_perc_t = FALSE`). The analytical $p$-value is also greater than 5%. For this example, bootstrap and normal approximation both do not reject. 

## License

This software is open source software and is licensed under the [GNU General Public License 3.0][gnu 3.0]. 

[gnu 3.0]: https://www.gnu.org/licenses/gpl-3.0.en.html
